The Chemistry Development Kit (CDK): atom typing, rendering, molecular formulas, and substructure searching

### 2016-09-25 article.tex
A number of small tweaks

\def \cdkversion {\textbf{2.0}}

@@

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\parttitle{Background}
Cheminformatics is a well-established field with many applications in chemistry,
biology, drug discovery, and others. The Chemistry Development Kit (CDK) has
become a widely used Open Source cheminformatics toolkit, providing various models to represent
chemical structures, of which the chemical graph is essential. However,
in the first five years of the project increased so much in size that
interdependencies between components grew unmanageable large, resulting
in unpredictable instabilities.
\parttitle{Results}
We here report improvements to the CDK since the 1.2 release series made
to accommodate both the increased complexity of the library, as well as
significant improvements of and additions to the functionality of the
library. Second, we outline how the CDK evolved with respect to
quality control and the approach we have adopted to ensure stability, including
a peer review mechanism.
Additionally, a selection of the new APIs that have been introduced will
be discussed: atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
\parttitle{Conclusions}
With this paper we have shown the continued effort to provide a free, Open Source
cheminformatics library, and show that such collaborative projects can exist over
a long period. We have taken advantage from the community support, and show
that an open source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

@@

    // lazily iterate over all unique atom matches as query atom
    // index bijection to atoms in 'mol', 'rxn' can also be used
    for (int[] m : pattern.matchAll(mol)
                          .uniqueAtoms()) {

    }
  \end{verbatim}

@@

CTfile Sgroups capture and organise high level information about sets of atoms
and bonds~\cite{Gushurst91}. There are four types of Sgroup: Display Shortcuts, Polymers,
Mixtures, and Data. The most familiar Sgroups from an end user perspective are structure 
repeat units (e.g. bracketing) and abbreviations (Figure~8\ref{fig:sgroups}).
CDK~\cdkversion{} adds 
supports for representation, reading, writing, and depiction.

@@

Since the last CDK publication in 2006, the library has been improved in many
aspects. The functionality has seen many additions and rewrites, making the
library more functional and a lot more faster then before. Updates on the common
SMILES and molfile formats and the improved structure diagram generation are
very visible and benefit many of the tools using the CDK.
Furthermore, the stability of the development model has significantly improved,
providing greater stability of the library over time. The project adopted peer
review and quality control.

With 93 contributors, a long list of tools based on the CDK, and hundreds
of article citations, the CDK is alive and kicking.

>>>>

%\def \cdkversion {\textbf{2.0}}
\def \cdkversion {2.0}

@@

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\parttitle{Background}
Cheminformatics is a well-established field with many applications in chemistry,
biology, drug discovery, and others. The Chemistry Development Kit (CDK) has
become a widely used Open Source cheminformatics toolkit, providing various models to represent
chemical structures, of which the chemical graph is essential. However,
in the first five years of the project increased so much in size that
interdependencies between components grew unmanageable large, resulting
in unpredictable instabilities.
\parttitle{Results}
We here report improvements to the CDK since the 1.2 release series made
to accommodate both the increased complexity of the library, as well as
significant improvements of and additions to the functionality of the
library. Second, we outline how the CDK evolved with respect to
quality control and the approach we have adopted to ensure stability, including
a peer review mechanism.
Additionally, a selection of the new APIs that have been introduced will
be discussed: atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
\parttitle{Conclusions}
With this paper we have shown the continued effort to provide a free, Open Source
cheminformatics library, and show that such collaborative projects can exist over
a long period and result in a high-quality and fast library. We have taken
advantage from the community support, and show
that an open source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

@@

    // lazily iterate over all unique atom matches as query atom
    // index bijection to atoms in 'mol', 'rxn' can also be used
    for (int[] m : pattern.matchAll(mol)
                          .uniqueAtoms()) {
       ...
    }
  \end{verbatim}

@@

CTfile Sgroups capture and organise high level information about sets of atoms
and bonds~\cite{Gushurst91}. There are four types of Sgroup: Display Shortcuts, Polymers,
Mixtures, and Data. The most familiar Sgroups from an end user perspective are structure 
repeat units (e.g. bracketing) and abbreviations (Figure~8\ref{fig:sgroups}).
CDK~\cdkversion{} adds 
supports for representation, reading, writing, and depiction of Sgroups.

@@

Since the last CDK publication in 2006, the library has been improved in many
aspects. The functionality has seen numerous additions and rewrites, making the
library more functional and a lot faster than before. Updates on the common
SMILES and molfile formats and the improved structure diagram generation are
very visible and benefit many of the tools using the CDK.
Furthermore, the stability of the development model has significantly improved,
providing greater stability of the library over time. The project adopted peer
review and quality control.

With more than 90 contributors, a long list of tools based on the CDK, and
hundreds of article citations, the CDK is alive and kicking.
